Compositional Certified Resource Bounds

### 2015-04-09 paper/main.tex
looks like PLDI conferenceinfo was broken
add nice and well deserved aec badge
add a note in the logical state section

\usepackage[utf8]{inputenc} %for utf8 input
\usepackage{amssymb} %for shift symbol
\usepackage{amsmath}
\usepackage{listings} %for code
\lstset{moredelim=[is][\color{blue}]{@}{@},keepspaces=true}
\usepackage{mathpartir} %for typing rules
\usepackage{microtype} %better micro typing
\usepackage{stmaryrd} %for llbracket
\usepackage{mathabx} % for boxes
\usepackage{graphicx} %to include png images
\usepackage{xcolor} %for colors
\usepackage{url}
\usepackage{enumitem}
\usepackage{array} %for stupid tables

@@

\conferenceinfo{PLDI'15}{, June 13--17, 2015, Portland, OR, USA}
\copyrightyear{2015}
\copyrightdata{978-1-4503-3468-6/15/06}
\doi{nnnnnnn.nnnnnnn}

@@

@@

More precisely, the technique consists of the following steps. First,
a program $P$ that fails to be analyzed automatically is enriched by
auxiliary variables $\vec x$ to form a program
$P_l(\vec x)$.  Second, an initial value $\vec X(\sigma)$ for the
logical variables is selected so that:
$$
\forall n\, \sigma\, \sigma'.\,
  (\sigma, P_l(\vec X(\sigma))) \Downarrow_n \sigma'
    {\implies}
  \exists n' {\le} n.\,
    (\sigma, P) \Downarrow_{n'} \sigma'.
$$
Since the annotated program and the original one are usually
syntactically close, the proof of this result goes
by simple induction on the resource-aware evaluation judgement.
Third, using existing automation tools, a bound $B(\vec x)$ for
$P_l(\vec x)$ is derived.  Finally this bound, instantiated with
$\vec X$, gives the final resource bound for the program $P$.

This idea is illustrated by the program in \pref{fig:xmplincaux}.
The parts of the code in blue are annotations that were added
to the original program text.  The top-level loop increments
a binary counter $k$ times.  A naive analysis of
the algorithm yields the quadratic bound $k \cdot N$.
However, the algorithm is in fact linear and its cost is bounded
by $2k + \#_1(a)$ where $\#_1(a)$ denotes
the number of one entries in the array $a$.  Since this
number depends on the heap contents, no tool available
for C is able to derive the linear bound.  However, it can be inferred
by our automated tool if a logical variable $na$ is introduced.
This logical variable is a reification of the
number $\#_1(a)$ in the program.  For example, on line 6 of the example
we are setting \lstinline{a[x]} to 0 and because of the condition we
know that this array entry was 1.  To reflect this change on
$\#_1(a)$, the logical variable $na$ is decremented.
Similarly, on line 8, an array entry which was 0 becomes 1, so
$na$ is incremented.  To complete the step 2 of the systematic
procedure described above, we must show that the extra assertion
\lstinline{na > 0} on line 5 is always true.  It
is done by proving inductively that $na = \#_1(a)$
and remarking that since \lstinline{a[x] == 1} is true, we must have
$\#_1(a) > 0$, thus the assertion \lstinline{na > 0} never fails. %  So the
% test is in fact semantically equivalent to \lstinline{true}.

>>>>

\usepackage[utf8]{inputenc} %for utf8 input
\usepackage{amssymb} %for shift symbol
\usepackage{amsmath}
\usepackage{listings} %for code
\lstset{moredelim=[is][\color{blue}]{@}{@},keepspaces=true}
\usepackage{mathpartir} %for typing rules
\usepackage{microtype} %better micro typing
\usepackage{stmaryrd} %for llbracket
\usepackage{mathabx} % for boxes
\usepackage{graphicx} %to include png images
\usepackage{xcolor} %for colors
\usepackage{url}
\usepackage{enumitem}
\usepackage{array} %for stupid tables
\usepackage[firstpage]{draftwatermark}
\SetWatermarkText{\hspace*{8in}\raisebox{5.5in}{\includegraphics[width=2.0cm]{fig/aec-badge-pldi.pdf}}}
\SetWatermarkAngle{0}

@@

\conferenceinfo{PLDI'15}{June 13--17, 2015, Portland, OR, USA}
\copyrightyear{2015}
\copyrightdata{978-1-4503-3468-6/15/06}
\doi{nnnnnnn.nnnnnnn}

@@

More precisely, the technique consists of the following steps. First,
a program $P$ that fails to be analyzed automatically is enriched by
auxiliary variables $\vec x$ and assertions to form a program
$P_l(\vec x)$.  Second, an initial value $\vec X(\sigma)$ for the
logical variables is selected so that:
$$
\forall n\, \sigma\, \sigma'.\,
  (\sigma, P_l(\vec X(\sigma))) \Downarrow_n \sigma'
    {\implies}
  \exists n' {\le} n.\,
    (\sigma, P) \Downarrow_{n'} \sigma'.
$$
Since the annotated program and the original one are usually
syntactically close, the proof of this result goes
by simple induction on the resource-aware evaluation judgement.
Third, using existing automation tools, a bound $B(\vec x)$ for
$P_l(\vec x)$ is derived.  Finally this bound, instantiated with
$\vec X$, gives the final resource bound for the program $P$.

This idea is illustrated by the program in \pref{fig:xmplincaux}.
The parts of the code in blue are annotations that were added
to the original program text.  The top-level loop increments
a binary counter $k$ times.  A naive analysis of
the algorithm yields the quadratic bound $k \cdot N$.
However, the algorithm is in fact linear and its cost is bounded
by $2k + \#_1(a)$ where $\#_1(a)$ denotes
the number of one entries in the array $a$.  Since this
number depends on the heap contents, no tool available
for C is able to derive the linear bound.  However, it can be inferred
by our automated tool if a logical variable $na$ is introduced.
This logical variable is a reification of the
number $\#_1(a)$ in the program.  For example, on line 6 of the example
we are setting \lstinline{a[x]} to 0 and because of the condition we
know that this array entry was 1.  To reflect this change on
$\#_1(a)$, the logical variable $na$ is decremented.
Similarly, on line 8, an array entry which was 0 becomes 1, so
$na$ is incremented.  To complete the step 2 of the systematic
procedure described above, we must show that the extra assertion
\lstinline{na > 0} on line 5 is always true.  It
is done by proving inductively that $na = \#_1(a)$
and remarking that since \lstinline{a[x] == 1} is true, we must have
$\#_1(a) > 0$, thus the assertion \lstinline{na > 0} never fails.

Generally, we observed that because the instrumented program
is structurally same as the original one, it is sufficient to prove
that the added assertions never fail to show that they have identical
resource consumption.  This can be piggybacked on standard
static-analysis tools.

@@
